HO unification from object language to meta language

### 2024-02-14 main.tex

\begin{elpicode}
  kind fo_tm type.
  type fo_app list fo_tm -> fo_tm.
  type fo_lam (tm -> fo_tm) -> fo_tm.
  type fo_uv  nat -> fo_tm.
  type fo_c string -> fo_tm.
\end{elpicode}

\begin{elpicode}
  kind tm type.
type app list tm -> tm.
type lam (tm -> tm) -> tm.
type uv  nat -> list tm -> tm.
type c string -> tm.
\end{elpicode}

@@

proprieta desiderate, serve equalho (fa deref)

>>>>

\begin{tabular}{c|c}
\begin{minipage}{.5\textwidth}
\begin{minted}{elpi}
kind fo_tm type.
type fo_app list fo_tm -> fo_tm.
type fo_lam (tm -> fo_tm) -> fo_tm.
type fo_uv  nat -> fo_tm.
type fo_c string -> fo_tm.

typeabbrev subst list (option tm).
type equal subst -> tm -> tm -> o.
\end{minted}
\end{minipage}
&
\begin{minipage}{.5\textwidth}
\begin{minted}{elpi}
kind tm type.
type app list tm -> tm.
type lam (tm -> tm) -> tm.
type uv  nat -> list tm -> tm.
type c string -> tm.

kind assmt type.
type abs (tm -> assmt) -> assmt.
type val tm -> assmt.

typeabbrev subst list (option assmt).
type equal subst -> tm -> tm -> o.
\end{minted}
\end{minipage}
\end{tabular}

@@

\begin{tabular}{c}
\begin{minipage}{.9\textwidth}
\begin{minted}{elpi}
% deref
equal S (uv N) T1 :- assigned? N S T, equal S T T1.
equal S T1 (uv N) :- assigned? N S T, equal S T1 T.
equal S (app [uv N|Args]) T1 :- assigned? N S F, beta F Args T, equal S T T1.
equal S T1 (app [uv N|Args]) :- assigned? N S F, beta F Args T, equal S T1 T.
% congruence
equal S (app L1) (app L2) :- forall2 (equal S) L1 L2.
equal S (lam F1) (lam F2) :- pi x\ equal S x x => equal S (F1 x) (F2 x).
equal _ (c X) (c X).
equal _ (uv N) (uv N).
% beta
equal S (app [lam X | TL]) T :- beta (lam X) TL T', equal S T' T.
equal S T (app [lam X | TL]) :- beta (lam X) TL T', equal S T T'.
% eta
equal S (lam F) T :- not (T = lam _),
  pi x\ beta T [x] (T' x), equal S (lam F) (lam T').
equal S T (lam F) :- not (T = lam _),
  pi x\ beta T [x] (T' x), equal S (lam T') (lam F).
\end{minted}
\end{minipage}
\\\hline
\begin{minipage}{.9\textwidth}
\begin{minted}{elpi}
% deref
equal S (uv N Args) T1 :- assigned? N S F, beta-abs F Args T, equal S T T1.
equal S T1 (uv N Args) :- assigned? N S F, beta-abs F Args T, equal S T1 T.
% congruence
equal S (app L1) (app L2) :- forall2 (equal S) L1 L2.
equal S (lam F1) (lam F2) :- pi x\ equal S x x => equal S (F1 x) (F2 x).
equal _ (c X) (c X).
equal _ (uv N A1) (uv N A2) :- forall2 (equal S) A1 A2. %fixme in the code
\end{minted}
\end{minipage}

\end{tabular}

equal fo is congruence, deref, beta, eta (for the OL).
equal ho is congruence, deref (for the OL).
unif (fo or ho) should cover pattern fragment, i.e. well behaved unif.

ho, and unif ho, is the LF/ML language/unif. here we describe fo and ho
with a deeply embedded syntax, but the idea is that uv can be mapped to
an actual unif variable.

of course we could write a unif for the ho that had beta and eta,
but if we use the one of the LF it can't have eta beta of the OL baked in.
